\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\usepackage{booktabs}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{longtable}
\input{preamble-req.tex}

\newcommand{\intent}[1]{\textit{Design intent (not yet implemented):} #1}
\newcommand{\proof}[1]{\textbf{Proof shipped (Track~A):} #1}
\newcommand{\trackb}[1]{\textbf{Track~B (Linux production):} #1}

\title{Demiurge: A Phase-Aware, Cache-Locality-First Load Balancer\\for Inference Fleets\\[4pt]
  \large Technical Design Specification v1.5}
\author{Demiurge Architecture Group}
\date{}

\begin{document}
\maketitle

\begin{abstract}
An inference request is a \textbf{lease on stateful accelerator memory}, not a
packet. Demiurge targets fleets that disaggregate \emph{prefill} (compute-bound,
cache-producing) from \emph{decode} (bandwidth-bound, cache-consuming), with the
KV cache as the explicit hand-off artifact. This document is the \textbf{target
design}; shipped scope is enumerated in \S\ref{sec:status}. Paragraphs marked
\intent{...} describe behavior planned in \texttt{ROADMAP.md}, not code that
exists today. Per-requirement \emph{implemented} vs \emph{intended} status lives
only in the generated conformance matrix (Table~\ref{tab:conformance}); normative
prose uses status-neutral requirement tags (the \texttt{\textbackslash req\{...\}} macro).
\end{abstract}

\section{Implementation Status}\label{sec:status}
\textbf{Track~A proof shipped (Phases~0--5).} Everything below runs on macOS and
Linux today with mock TCP backends, userspace dataplane proofs, and CI gates.
\begin{itemize}[nosep]
  \item \textbf{Phase~0} --- Cost-factor algebra: log-space composition, fail-expensive
        clamping, bounded corrector (\req{DEMI-COST-POS}, \req{DEMI-CORR-CLAMP},
        \req{DEMI-FAIL-EXPENSIVE}); minimum-cost backend selection over phase pools
        (\req{DEMI-ROUTE-MINCOST}).
  \item \textbf{Phase~1} --- Non-blocking \texttt{Route}: async prefill dispatch,
        decode continuation on \texttt{OnPrefillComplete} (\req{ALG-ROUTE});
        short-context colocated fast path (\req{DEMI-SHORT-FASTPATH}).
  \item \textbf{Phase~2} --- KV hand-off over a TCP proof transport
        (\req{DEMI-KV-HANDOFF}, \req{DEMI-HANDOFF-XPORT}); overhead-aware reservation
        and fleet $\Phi$ barrier (\req{DEMI-KV-OVERHEAD}, \req{DEMI-BARRIER-PHI},
        \req{DEMI-KV-RELEASE}); hand-off transfer telemetry
        (\req{DEMI-KV-TRANSFER-TELEM}).
  \item \textbf{Phase~3} --- AP state plane with warmth maps and gossip merge
        (\req{DEMI-WARM-DISCOUNT}, \req{DEMI-STATE-AP}); fast-path warmth override.
  \item \textbf{Phase~4} --- Greedy pf$\to$dc pairing (\req{DEMI-PAIR-GREEDY});
        shadow pool rebalancer (\req{DEMI-POOL-RATIO}); fleet shadow replay
        (\req{DEMI-FLEET-SHADOW}); trace-driven fleet simulation spinoff
        (\req{DEMI-FLEET-SIM}); fast-path telemetry export
        (\req{DEMI-FASTPATH-TELEM}); RDMA transfer cost shadow
        (\req{DEMI-RDMA-COST-SHADOW}); corrector shadow logging
        (\req{DEMI-CORR-SHADOW}).
  \item \textbf{Phase~5 (proof)} --- RCU routing snapshot on the live TCP path
        (\req{DEMI-DP-RCU}); userspace admit-bucket shed before L7
        (\req{DEMI-XDP-SHED}).
\end{itemize}

\textbf{Track~B engineering (Linux).} Kernel dataplane code ships and passes
\texttt{./scripts/track-b-verify.sh} on Linux VM (veth smoke; production NIC exit
gates remain open):
\begin{itemize}[nosep]
  \item Runtime XDP attach + kernel admit (\texttt{AdmitMode}, actuation BPF map sync).
  \item \texttt{io\_uring} L7 forwarder on production TCP \texttt{serve()}
        (\texttt{IoUringProxySession} recv/send per connection).
  \item Track~B load scenarios (\texttt{LOAD-TRACK-B-IOURING}, \texttt{LOAD-TRACK-B-KERNEL}).
\end{itemize}

\textbf{Not shipped (Track~B exit gates / Track~C).} XDP shed on production NIC
under fleet load, data-plane p99 admit budget on x86\_64 reference hardware under
CP slowdown, pool rebalancing \emph{actuation} in production fleets, RDMA production
hand-off, cross-tenant sharing (\S\ref{sec:sharing}), live migration
(\S\ref{sec:migration}), and corrector graduation (\S\ref{sec:corrector}) remain
\textbf{design intent} tracked in \texttt{design/requirements.toml} and
\texttt{ROADMAP.md}.

\begin{table}[h]
  \centering
  \input{generated/conformance_matrix.tex}
  \caption{Conformance matrix (generated from \texttt{design/requirements.toml}).
  \emph{Implemented} rows are backed by named tests in CI; \emph{intended} rows
  are design targets only. Never hand-edit this table.}
  \label{tab:conformance}
\end{table}

\section{Document Model and Traceability}\label{sec:docmodel}
Three artifacts stay strictly separated so target design never blurs with
shipped scope:
\begin{enumerate}[nosep]
  \item \textbf{This spec} (\texttt{spec/demiurge.tex}) --- \emph{what} the system
        should do. Normative claims carry requirement IDs via \texttt{\textbackslash req\{...\}};
        unimplemented behavior
        is wrapped in \intent{...}.
  \item \textbf{Requirements registry} (\texttt{design/requirements.toml}) ---
        \emph{whether} each registered ID is \texttt{implemented} (code + named tests)
        or \texttt{intended} (design only), plus development phase.
  \item \textbf{Generated tables} (\texttt{spec/generated/}) --- read-only
        projections of canonical inputs; CI fails on drift.
\end{enumerate}

\textbf{Blur guard.} \texttt{cargo xtask lint} enforces a three-way join
(spec~$\leftrightarrow$~code~$\leftrightarrow$~tests) \emph{and} placement rules:
\begin{itemize}[nosep]
  \item \texttt{intended} requirements: every tagged ID in this document must appear
        \emph{only} inside \intent{...} paragraphs.
  \item \texttt{implemented} requirements: every tagged ID must appear at least once
        \emph{outside} \intent{...} (normative or status summary prose).
\end{itemize}
Status strings never appear in hand-authored spec prose; only the generated matrix
carries \texttt{implemented}/\texttt{intended} columns.

\textbf{Parameters.} Tunable constants live in \texttt{design/demiurge.params.toml};
\texttt{cargo xtask gen} projects them into Rust constants and
Table~\ref{tab:params}. CPU hot-path gates use the same pattern via
\texttt{design/bench-gates.toml}.

\section{Development Tracks}\label{sec:tracks}
Phases (0--8) order \emph{dependencies}; tracks order \emph{where code runs}.
Requirement phase numbers are stable; burndown is reported by
\texttt{cargo xtask lint} (\texttt{P0: 4/4}, \ldots).

\begin{table}[h]
  \centering
  \small
  \begin{tabular}{lllp{6.5cm}}
    \toprule
    Track & Platform & Phases & Scope \\
    \midrule
    A & macOS / portable Rust & 0--5 proof & Mock TCP, userspace RCU/admit, full CI \\
    A+ & macOS + trace replay & 4, 8 shadow & Fleet pilot, corrector shadow, HandoffTransport \\
    B & Linux x86\_64 & 5+ prod & XDP compile/attach, \texttt{io\_uring} forwarder \\
    C & Linux + GPU fleet & 6--8 & Migration, sharing auth, corrector graduation \\
    \bottomrule
  \end{tabular}
  \caption{Development tracks (see \texttt{ROADMAP.md} for gates and CI wiring).}
  \label{tab:tracks}
\end{table}

\proof{\texttt{DEMI-DP-RCU} and \texttt{DEMI-XDP-SHED} at \emph{implemented}: RCU
snapshot read on the live TCP path; userspace \texttt{AdmitBucket} shed before L7.
Kernel XDP shed on a production NIC under fleet load is a Track~B exit gate.}
\trackb{Runtime XDP on veth, kernel admit wiring, and \texttt{io\_uring}
production TCP \texttt{serve()} are shipped; exit gates (production NIC shed,
x86\_64 p99 budget) remain on reference hardware.}

\section{Architecture}\label{sec:arch}
Three planes, three consistency models:
\begin{itemize}[nosep]
  \item \textbf{Data plane} (microsecond): eBPF/XDP admission, Rust
        \texttt{io\_uring} L7 forwarder, RCU routing snapshot
        (\S\ref{sec:dataplane}).
  \item \textbf{Control plane} (millisecond): cost evaluation, length
        prediction, pool-ratio autoscaling, CP consensus (\S\ref{sec:control}).
  \item \textbf{State plane} (AP): KV warmth map, live occupancy gossip
        (\S\ref{sec:state}). A wrong warmth guess costs a cache miss, never a
        correctness violation.
\end{itemize}
\intent{Authorization for cache sharing stays strongly consistent on the control
plane (\S\ref{sec:sharing}); cache \emph{warmth} stays AP.}

\section{Cost Function}\label{sec:cost}
The analytic routing cost is a product of a strictly-positive time core and
dimensionless factors:
\[
  C(r,q) \;=\; T_{\text{core}}(r,q)\;\cdot\;\prod_m B_m\;\cdot\;\prod_c D_c\;\cdot\;\delta,
  \qquad B_m \ge 1,\; D_c \in (0,1],\; \delta \in [1-\alpha,\,1+\alpha].
\]

\subsection{Positivity}\label{sec:cost-pos}
Because $T_{\text{core}}>0$ and every factor is strictly positive, $C(r,q)>0$ by
construction \req{DEMI-COST-POS}. The implementation stores $\ln C$ and composes
by adding logs, so the guarantee survives floating point: comparisons use the
exact log; there is no linear product to underflow to zero or flip sign.

\subsection{Composition and Hot-Path Signals}\label{sec:cost-compose}
Rewards enter only as multiplicative discounts $D_c\in(0,1]$, never as
subtractions. Invalid or non-finite hot-path signals saturate \emph{toward
expensive} \req{DEMI-FAIL-EXPENSIVE}, so a broken metric cannot cheapen a
target. Canonical $\alpha$ and related constants are generated from
\texttt{design/demiurge.params.toml} (Table~\ref{tab:params}).

\begin{table}[h]
  \centering
  \input{generated/params_table.tex}
  \caption{Canonical parameters (generated). Groups for future phases are
  placeholders until their roadmap phase lands.}
  \label{tab:params}
\end{table}

\section{KV Cache Overhead Accounting}\label{sec:kv}
Per-request reservation uses block-aligned token counts plus metadata and
fragmentation slack (parameters \texttt{kv.*}). The $\Phi$ memory-pressure
barrier uses \textbf{fleet-aggregate} occupancy, not a sum of per-request p90
headroom \req{DEMI-BARRIER-PHI}. Reservations are released on session end, abort,
or TTL expiry \req{DEMI-KV-RELEASE}. Overhead terms are mandatory in the
reservation formula \req{DEMI-KV-OVERHEAD}.

\section{Cross-Tenant Cache Sharing}\label{sec:sharing}
\intent{Prefix-cache reuse across tenants is opt-in via a Shared-Prefix Group
with a content-verified template. Cache-domain keys are tenant-salted; a
non-member never resolves to a shared cache-domain key \req{DEMI-S1-DOMAIN}.}

\subsection{Shared-Domain Isolation}\label{sec:sharing-s1}
\intent{Membership is checked on the strongly consistent authorization path before
any warmth discount is applied.}

\section{State Plane}\label{sec:state}
Each backend publishes a block-granularity warmth map and live occupancy via
gossip (\texttt{telemetry.*}). A warmth hit applies a bounded routing discount
\req{DEMI-WARM-DISCOUNT}; a miss applies none. The plane is AP: stale warmth
causes miss only \req{DEMI-STATE-AP}.

\section{Learned Corrector}\label{sec:corrector}
The analytic corrector multiplier $\delta$ is clamped to $[1-\alpha,1+\alpha]$
\req{DEMI-CORR-CLAMP}. Shadow mode logs $(\text{features}, \ln C, t_{\mathrm{obs}})$
and trains a bounded $\delta$ offline without actuation \req{DEMI-CORR-SHADOW}.
\intent{Production use of a learned $\delta$ follows a
shadow $\to$ canary $\to$ prod graduation path; promotion requires no clamp
saturation and no violation of \req{DEMI-COST-POS} \req{DEMI-CORR-GRAD}. Until
Phase~8, the router uses $\delta=1$.}

\section{Routing and Migration}\label{sec:routing}

\subsection{Minimum-Cost Selection}\label{sec:route-mincost}
Within a phase pool, the forwarder selects the backend with minimum $\ln C$
\req{DEMI-ROUTE-MINCOST}. Live in-flight load enters via a queueing barrier on
\texttt{Backend::cost()}.

\subsection{Non-Blocking Route and Fast Path}\label{sec:route}
\texttt{Route} admits a request, dispatches prefill asynchronously, and
returns immediately \req{ALG-ROUTE}; decode placement runs in
\textsc{OnPrefillComplete} on fresh telemetry. Prompts at or below
\texttt{routing.short\_context\_tokens} use a \textbf{colocated} fast path
(single backend, no cross-pool hand-off) unless a warmth override forces
disaggregation \req{DEMI-SHORT-FASTPATH}.

\begin{algorithm}[h]
\caption{Route (target behavior; Phase~1)}
\begin{algorithmic}[1]
\State \textbf{on} request $r$:
\State classify: fast path \textbf{or} disaggregated
\State \intent{admit under tenant bucket or shed at XDP (\S\ref{sec:dataplane})}
\State $p \gets \textsc{SelectPrefillTarget}(r)$
\State \textbf{dispatch} prefill of $r$ to $p$ \textbf{asynchronously}
\State \textbf{return} \Comment{decode in \textsc{OnPrefillComplete}}
\end{algorithmic}
\end{algorithm}

\subsection{KV Hand-Off}\label{sec:handoff}
Prefill completion publishes a KV hand-off handle; decode placement never
proceeds without it \req{DEMI-KV-HANDOFF}. Each completed transfer records byte
length and prefill I/O wall time; operators report p50/p99 aggregates
\req{DEMI-KV-TRANSFER-TELEM}. Transport is pluggable via
\texttt{HandoffTransport} (TCP header pass-through default; mock RDMA for local
proof) \req{DEMI-HANDOFF-XPORT}. Disaggregated pf$\to$dc placement logs a
shadow RDMA transfer model (topology distance + bytes/bandwidth) alongside the
flat cross-node penalty; routing still uses the flat barrier
\req{DEMI-RDMA-COST-SHADOW}.

\subsection{Greedy Pairing}\label{sec:pairing}
Joint pf$\to$dc placement is approximated greedily: fix prefill target,
then optimize decode conditional on it \req{DEMI-PAIR-GREEDY}. Pairing regret is
measured in shadow.

\subsection{Live Migration}\label{sec:migration}
\intent{Decode migration is chunked and abortable. Cutover commits only if
estimated stall $\le \varepsilon\cdot\mathrm{ITL}$ \req{DEMI-MIG-SUBITL};
otherwise the original placement is restored.}

\section{Control Plane}\label{sec:control}
The pool-ratio controller adjusts prefill vs decode capacity share
$\pi$ using queue, KV, and SLO pressure with hysteresis and cooldown
(\texttt{pool.*}) \req{DEMI-POOL-RATIO}. Shadow mode is default until Phase~5
actuation. Shadow fleet replay on a held-out trace must show $\pi^*$ tracking
prefill-heavy windows \req{DEMI-FLEET-SHADOW}. The \emph{'sim} spinoff replays
the same trace windows as live TCP load against mock backends, with per-window
knobs and fleet gates \req{DEMI-FLEET-SIM}. The router exports
\texttt{fast\_path\_ratio} and near-threshold mis-route regret
\req{DEMI-FASTPATH-TELEM}.

\section{Data Plane}\label{sec:dataplane}
The live TCP forwarder reads the last published RCU routing snapshot and never
blocks on control-plane publish \req{DEMI-DP-RCU}. Overload shedding uses a
token-bucket admit gate before L7 \req{DEMI-XDP-SHED}.

\begin{table}[h]
  \centering
  \small
  \begin{tabular}{lll}
    \toprule
    Capability & Track~A proof & Track~B production \\
    \midrule
    RCU snapshot read & Userspace \texttt{DataPlaneSnapshot} & Same + pinned reader threads \\
    Admit / shed & \texttt{AdmitBucket} in router & eBPF/XDP map + token bucket \\
    L7 proxy & \texttt{TcpStream} copy & \texttt{io\_uring} registered buffers \\
    \bottomrule
  \end{tabular}
  \caption{Dataplane proof vs production (honest scope for Phase~5).}
  \label{tab:dataplane}
\end{table}

\trackb{Production wiring attaches compiled BPF to netdev (veth smoke shipped;
production NIC exit gate open), programs bucket capacity from control-plane $\pi$,
and routes bulk bytes through \texttt{io\_uring} on the production TCP
\texttt{serve()} loop (\texttt{IoUringProxySession}).}

\section{Tooling, Gates, and Operations}\label{sec:ops}
Design conformance is enforced in CI and locally via \texttt{./scripts/gate.sh}:
\begin{itemize}[nosep]
  \item \texttt{cargo xtask gen} --- regenerate params, conformance matrix, Rust constants.
  \item \texttt{cargo xtask lint} --- traceability join, phase burndown, blur guard.
  \item \texttt{cargo xtask bench-gate} --- release-mode CPU gates (\texttt{bench-gates.toml}).
  \item \texttt{cargo xtask spec} --- \texttt{gen} + \LaTeX{} build $\to$ \texttt{spec/demiurge.pdf}.
  \item \texttt{./scripts/load-bench.sh} / \texttt{load-stress.sh} --- TCP load scenarios (stress local-only).
  \item \texttt{cargo xtask fleet-pilot} --- shadow fleet replay (Track~A+).
  \item \texttt{cargo xtask apostrophe-sim} --- trace $\to$ live load fleet simulation \req{DEMI-FLEET-SIM}.
\end{itemize}

\textbf{Operational posture.} The data plane is fail-closed on admit exhaustion
(503 before backend work). Invalid cost signals fail expensive, never cheap.
Warmth and occupancy are AP: stale state causes cache miss, not policy violation.
\intent{Cross-tenant cache authorization remains CP-strong (\S\ref{sec:sharing}).}

\textbf{Release artifacts.} Tagged macOS releases via \texttt{scripts/publish.sh};
Linux \texttt{linux-nightly} binaries include BPF objects. Spec PDF is built on
every \texttt{spec/**} change in CI and uploaded as an artifact.

\section{Requirement Lifecycle}\label{sec:lifecycle}
New normative behavior follows the same PR pattern:
\begin{enumerate}[nosep]
  \item Add a requirement ID to spec prose with \texttt{\textbackslash req\{DEMI-...\}}
        (\intent{...} if not shipping yet).
  \item Register in \texttt{design/requirements.toml} with \texttt{status},
        \texttt{phase}, and backing \texttt{tests} when implemented.
  \item Reference the same ID in Rust doc comments and name \texttt{\#[test]} functions in the registry.
  \item Run \texttt{cargo xtask gen \&\& cargo xtask lint}; commit generated files.
\end{enumerate}
Phases close only by flipping \texttt{intended} $\to$ \texttt{implemented} with
named tests --- never the reverse (ratchet only tighter).

\end{document}
