\documentclass[11pt,a4paper]{article}

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{amsmath,amssymb,amsthm}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{microtype}
\usepackage{enumitem}

\hypersetup{
  colorlinks=true,
  linkcolor=black,
  citecolor=black,
  urlcolor=black
}

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}{Definition}
\newtheorem{property}{Property}

\title{%
  \textbf{Log-Space Cost Composition}\\
  \large The Routing Algebra of Demiurge
}
\author{Demiurge Project\\\small\texttt{fxdv/demiurge}}
\date{June 2026}

\begin{document}
\maketitle

\begin{abstract}
Demiurge routes LLM inference requests by minimum analytic cost over phase pools
(prefill, decode) under cache-locality and memory-pressure constraints.
This note formalizes the \emph{cost-factor algebra} implemented in
\texttt{demiurge-cost} and the \emph{decision-makers} in router and control
planes that consume it: admission, path classification, min-cost placement,
greedy pairing, KV reservation, and pool-ratio actuation.
The cost $C(r,q) > 0$ is built as a product of a time core and dimensionless factors,
realized in software as \emph{addition in log-space}.
We state the algebraic domains of barriers, discounts, and correctors;
prove positivity and monotonicity under composition; characterize the
\emph{fail-expensive} clamping policy for invalid telemetry; and extend the
core with fleet $\Phi$ barriers, warmth discounts, cross-node transfer penalties,
and pool-ratio scaling.
The algebra is executable: property tests and CPU bench gates enforce the
invariants on every code merge (CI gate).
\end{abstract}

\section{Introduction}

Classical load balancers minimize connection counts or queue length.
For disaggregated LLM inference, that objective is mis-specified: the valuable
state is the \textbf{KV cache}, occupancy is stochastic, and the marginal cost
of placing request~$r$ on backend~$q$ depends on $q$'s current batch, warmth,
and fleet-level memory headroom.

Demiurge therefore evaluates a \textbf{strictly positive} routing cost
$C(r,q)$ and selects $\arg\min_q \ln C(r,q)$.
Positivity is not an after-the-fact assertion---it is a \textbf{representation
invariant}: costs live in log-space, composition is log-addition, and invalid
signals can only move a target toward \emph{more} expensive, never cheaper.

This document is the mathematical companion to
\texttt{spec/demiurge.tex}~\S4 and the reference implementation
\texttt{crates/demiurge-cost}.
Requirement IDs (\texttt{DEMI-COST-POS}, etc.) tie claims to tests.

\section{Preliminaries}

Fix a request context $r$ (tokens, tenant, routing key blocks) and a
candidate backend $q$ in a phase pool.
Let $\mathbb{R}_{++} = (0,\infty)$.

\begin{definition}[Log-cost representation]
A \emph{cost} is a finite real $\ell = \ln C$ with $C \in \mathbb{R}_{++}$.
Comparison uses $\ell$ only; linear values $C = e^{\ell}$ are for display.
\end{definition}

We write $\oplus$ for log-space composition (ordinary addition on logs):
\[
  \ell_1 \oplus \ell_2 \;=\; \ell_1 + \ell_2,
  \qquad
  C_1 \otimes C_2 \;=\; e^{\ell_1 + \ell_2}.
\]

\section{Factor algebra}

The analytic cost is a multiplicative product in linear space, subject to
the domain constraints below:
\begin{equation}\label{eq:product}
  C(r,q)
  \;=\;
  T_{\mathrm{core}}(r,q)
  \;\cdot\;
  \prod_{m} B_m
  \;\cdot\;
  \prod_{c} D_c
  \;\cdot\;
  \delta,
\end{equation}
In implementation, Equation~\eqref{eq:product} becomes
\begin{equation}\label{eq:logcompose}
  \ln C
  \;=\;
  \ln T_{\mathrm{core}}
  \;+\;
  \sum_m \ln B_m
  \;+\;
  \sum_c \ln D_c
  \;+\;
  \ln \delta,
\end{equation}
with the corrector term omitted when $\delta = 1$ (identity).

\subsection{Primitive factors}

\begin{definition}[Time core]
$T_{\mathrm{core}} \in \mathbb{R}_{++}$ is the dimensional anchor (seconds).
Constructor \texttt{TimeCore} rejects non-finite and non-positive values.
\end{definition}

\begin{definition}[Barrier]
$B \in [1, \infty)$ is a multiplicative \emph{penalty}.
Invalid input clamps to $B = \texttt{f64::MAX}$ (fail-expensive).
\end{definition}

\begin{definition}[Discount]
$D \in (0, 1]$ is a multiplicative \emph{reward}; it can only decrease cost.
Invalid input clamps to $D = 1$ (neutral---no unearned reward).
\end{definition}

\begin{definition}[Corrector]
$\delta \in [1-\alpha,\, 1+\alpha]$ with canonical $\alpha = 0.20$.
Production routing currently uses $\delta = 1$; shadow training estimates
bounded $\delta$ offline.
\end{definition}

The sole constructor \texttt{compose(core, barriers, discounts, corrector)}
returns \texttt{Cost\{ ln \}}.

\subsection{Composition operator}

\begin{theorem}[Strict positivity]\label{thm:pos}
If $T_{\mathrm{core}} > 0$, each $B_m \ge 1$, each $D_c \in (0,1]$, and
$\delta > 0$, then $C(r,q) > 0$.
Moreover, if all factor logs are finite, $\ln C$ is finite.
\end{theorem}

\begin{proof}
Every factor is strictly positive by domain.
Products of strictly positive reals are strictly positive.
Log-space: finite sum of finite terms; $e^{\ln C} > 0$.
\end{proof}

\begin{property}[No underflow / sign flip]
Unlike a linear product $\prod_i x_i$ with many small discounts,
log-addition cannot underflow to $0$ or change sign when discounts
approach~$0^+$.
Property tests sweep $n \in [1,256]$ factors at \texttt{f64::MIN\_POSITIVE}
(\texttt{cost\_log\_is\_finite\_at\_extremes}).
\end{property}

\begin{theorem}[Monotonicity of barriers and discounts]
Holding other factors fixed:
(i)~increasing any $B_m$ increases $\ln C$;
(ii)~decreasing any $D_c$ (stronger reward) decreases $\ln C$.
\end{theorem}

\begin{proof}
$\partial \ln C / \partial \ln B_m = 1 > 0$;
$\partial \ln C / \partial \ln D_c = 1$ with $D_c \le 1$.
\end{proof}

\section{Fail-expensive semantics}

Telemetry on the microsecond routing path can be missing or corrupt.
Demiurge adopts a \textbf{fail-expensive} policy: invalid signals never
\emph{cheapen} a target.

\begin{definition}[Fail-expensive clamp]
Let $\mathrm{clamp}_T(x)$ map invalid $x$ to $T_{\max}$ (the largest finite
representable time, e.g., \texttt{f64::MAX} seconds).
Let $\mathrm{clamp}_B(x)$ map invalid $x$ to $B_{\max}$.
Let $\mathrm{clamp}_D(x)$ map invalid $x$ to $1$.
\end{definition}

\begin{theorem}[Invalid signals never cheapen]\label{thm:fail}
For any valid baseline cost $C_0$ built from valid factors:
\[
  C(\mathrm{clamp}_T(\mathrm{NaN}), \ldots) \;\ge\; C_0,
  \qquad
  C(\ldots, \mathrm{clamp}_D(\mathrm{NaN}), \ldots) \;\ge\; C_0,
\]
when the invalid factor is the only perturbation.
\end{theorem}

\begin{proof}
If the baseline $C_0$ contains a valid discount $D_c \in (0,1)$, clamping
the invalid telemetry to $1$ removes this discount, yielding
$C = C_0 / D_c \ge C_0$. Equality holds only if the baseline factor was
already $1$.
\end{proof}

Violations increment \texttt{FACTOR\_CLAMP\_EVENTS} (process-global health gauge).

Barrier assembly into the fixed stack used by \texttt{service\_cost} follows
the same policy: overflow replaces the last slot with $B_{\max}$ rather than
silently dropping a penalty.

\paragraph{Cost vs admission.}
The algebra \emph{ranks} placement candidates. Hard KV and connection
capacity are enforced by the reservation ledger and admit bucket.
The $\Phi$ barrier soft-caps utilization at $0.999$ so $B_\Phi$ remains
finite; cost alone is not a hard reject wall.

\paragraph{Warmth envelope.}
The maximum warmth discount $\rho_{\max}$ (\texttt{warmth.max\_discount})
is a first-class parameter: at full hit it can dominate the corrector
band $\alpha$. Forged warmth is a trust-boundary concern (threat model T1/T3),
not merely a tuning knob.

\paragraph{Observed vs claimed $T_{\mathrm{core}}$.}
Live routing uses $\max(T_{\mathrm{claimed}},\, T_{\mathrm{ewma}})$ where
$T_{\mathrm{ewma}}$ tracks measured prefill wall, so under-reported claims
cannot undercut observation (threat model T6).

\section{Corrector envelope}

The learned multiplier $\delta$ adjusts analytic cost within a fixed band.

\begin{theorem}[Corrector bound]\label{thm:corr}
For any valid composition with identity corrector $\delta = 1$ and clamped
$\delta' \in [1-\alpha, 1+\alpha]$,
\[
  \bigl|\ln C' - \ln C\bigr|
  \;\le\;
  \max\bigl\{|\ln(1+\alpha)|,\, |\ln(1-\alpha)|\bigr\}.
\]
\end{theorem}

Thus the corrector cannot arbitrarily rescale cost; it operates within a
multiplicative envelope independent of fleet scale.
With $\alpha = 0.20$, the relative adjustment is at most $\pm 20\%$.

\section{Domain extensions}

The core algebra is extended by \emph{derived barriers and discounts} that
remain within the same composition law~\eqref{eq:logcompose}.

\subsection{Queue pressure}

Live in-flight load on backend~$q$ enters as
\begin{equation}
  B_{\mathrm{queue}} = 1 + n_{\mathrm{inflight}}(q).
\end{equation}
This is a barrier ($\ge 1$) and is monotonic in occupancy.

\subsection{Fleet $\Phi$ memory barrier}

KV reservation uses block-aligned accounting with metadata and fragmentation:
\begin{align*}
  \tau &= \left\lceil \frac{\text{prompt tokens}}{\text{block}} \right\rceil \cdot \text{block}, \\
  \text{payload} &= \tau \cdot \text{bytes\_per\_token}, \\
  \text{reserved} &= \text{payload} \cdot (1 + \eta_{\mathrm{meta}} + \eta_{\mathrm{frag}}).
\end{align*}

Fleet pressure uses \textbf{aggregate} occupancy, not a sum of per-request
p90 headroom (requirement \texttt{DEMI-BARRIER-PHI}):
\[
  F = F_{\mathrm{live}} + \Delta p_{90},
  \qquad
  \text{not } F_{\mathrm{live}} + N \cdot \Delta p_{90}.
\]

Utilization $u = F / \text{capacity} \in [0,1)$ yields the barrier
\begin{equation}\label{eq:phi}
  B_\Phi(u) = \frac{1}{1-u},
\end{equation}
monotone in $u$, with $B_\Phi \to \infty$ as $u \to 1^{-}$ (hard memory wall).

\subsection{Warmth discount}

A warmth hit of strength $s \in (0,1]$ maps to
\begin{equation}
  D_{\mathrm{warm}}(s) = 1 - s \cdot \rho_{\max},
  \qquad
  \rho_{\max} = 0.85.
\end{equation}
Miss ($s \le 0$) applies \emph{no} discount---stale AP state fails toward
neutral, not toward cheap.

\subsection{Cross-node transfer}

Disaggregated prefill$\to$decode placement with distinct labels incurs
\begin{equation}
  B_{\mathrm{xfer}} = \max(1,\, \pi_{\mathrm{transfer}}),
  \qquad
  \pi_{\mathrm{transfer}} = 1.05 \;\text{(canonical)}.
\end{equation}

A \textbf{shadow} RDMA model logs analytic transfer time
\[
  t_{\mathrm{rdma}} = t_{\mathrm{base}}(\mathrm{distance}) + \frac{\text{bytes}}{\text{bandwidth}},
\]
with discrete topology distance (node $<$ rack $<$ cluster $<$ cross-cluster).
Routing still uses $B_{\mathrm{xfer}}$; shadow $\ln t_{\mathrm{rdma}}$ is
telemetry only (\texttt{DEMI-RDMA-COST-SHADOW}).

\subsection{Pool-ratio scaling}

The control plane publishes prefill share $\pi \in (0,1)$.
Let $T_0$ denote the unscaled time core returned by the ordinary composition
(before pool-ratio modulation).
Service time cores scale inversely with pool pressure:
\begin{equation}
  T_{\mathrm{core}}^{\mathrm{pf}} = T_0 \cdot \frac{1-\pi}{\pi},
  \qquad
  T_{\mathrm{core}}^{\mathrm{dc}} = T_0 \cdot \frac{\pi}{1-\pi}.
\end{equation}
High $\pi$ (prefill-heavy fleet) raises decode-relative cost and lowers
prefill-relative cost---biasing min-cost selection without breaking
positivity.

\section{Decision-makers}

The cost algebra answers \emph{how expensive} a candidate placement is.
\textbf{Decision-makers} are the discrete policies that consume $\ln C$ and
commit actions: admit or shed, colocate or disaggregate, pick a backend,
reserve KV bytes, or republish fleet ratio~$\pi$.
They are grouped by plane and latency budget (Table~\ref{tab:deciders}).

\begin{table}[t]
  \centering
  \caption{Decision-makers and their objectives}
  \label{tab:deciders}
  \small
  \begin{tabular}{@{}p{2.1cm}p{2.4cm}p{2.6cm}p{4.8cm}@{}}
    \toprule
    Agent & Plane & Trigger & Decision rule \\
    \midrule
    Admit/Shed & Data & Connection arrival & Token bucket: pass or \texttt{503} \\
    Path classifier & Router & HTTP head parsed & Colocated, disaggregated, or decode-only \\
    \textsc{SelectPrefill} & Router & Disaggregated route & $\arg\min_p \ln C_{\mathrm{pf}}$ \\
    \textsc{SelectDecode} & Router & Prefill complete & $\arg\min_d \ln C_{\mathrm{dc}}(\cdot \mid p)$ \\
    KV ledger & Control & Decode admit & Reserve if $F + \Delta p_{90} \le \text{cap}$; else reject \\
    Pool rebalancer & Control & Periodic / trace & Publish $\pi$ if $|\pi^* - \pi| > \varepsilon$ and cooldown \\
    Length predictor & Control & Observed prompts & Fleet $p90$ increment for $\Phi$ (not direct pick) \\
    Migration gate & Control & Cutover attempt & Commit iff $\widehat{t}_{\mathrm{stall}} \le \varepsilon \cdot \mathrm{ITL}$ \\
    Corrector trainer & Control (shadow) & Offline batch & $\delta \in [1\!\pm\!\alpha]$; no live actuation yet \\
    \bottomrule
  \end{tabular}
\end{table}

Decision-makers that \emph{optimize} placement share one rule: compare
\textbf{only} $\ln C$ (never linear $C$). Hard gates (admit shed, KV reject,
migration abort) sit \emph{outside} min-cost and fail closed.

\subsection{Admission and overload shed}

Before any routing score is evaluated, the data plane may shed load.
Userspace \texttt{AdmitBucket} and kernel XDP token buckets enforce a burst
capacity $B_{\mathrm{admit}}$; exhaustion returns HTTP~503 without entering
the routing loop.
When pool ratio~$\pi$ actuates, admit capacity rescales with $\pi$ so overload
protection tracks fleet pressure.
For instance, the token bucket refill rate for the prefill admission gate is
multiplied by $\pi$, and for the decode gate by $1-\pi$.
This is a \textbf{boolean gate}, not a min-cost problem---but it precedes every
cost-based decision-maker below.

\subsection{Path classifier}

\texttt{Route} classifies each request into a \textbf{path} before backend
selection (\texttt{ALG-ROUTE}):

\begin{enumerate}[nosep]
  \item \textbf{Decode-only} --- client declares decode phase; pick
        $\arg\min_{d} \ln C_{\mathrm{dc}}$ only.
  \item \textbf{Colocated fast path} --- prompt tokens
        $\le \tau_{\mathrm{short}}$ and no warmth override; pick
        $\arg\min_{q} \ln C_{\mathrm{colocated}}(r,q)$ on a single pool;
        no hand-off handle (\texttt{DEMI-SHORT-FASTPATH}).
  \item \textbf{Disaggregated} --- long context, warmth override, or unknown
        length (conservatively above $\tau_{\mathrm{short}}$); async prefill
        then decode in \textsc{OnPrefillComplete}.
\end{enumerate}

Warmth override fires when hit strength
$s \ge \texttt{routing.short\_context\_warmth\_override}$ on a remote prefill
target, forcing disaggregation despite a short prompt.

\subsection{Minimum-cost selection}

Within a phase pool,
\begin{equation}
  q^\star = \arg\min_{q \in \mathcal{Q}} \ln C(r,q).
\end{equation}
Implementation compares \texttt{Cost::ln()} only (\texttt{DEMI-ROUTE-MINCOST}).
Live inflight count enters via $B_{\mathrm{queue}} = 1 + n_{\mathrm{inflight}}$;
published $\pi$ scales $T_{\mathrm{core}}$ through \texttt{pool\_core\_scale}.

\subsection{Greedy pairing}

Joint prefill$\to$decode placement is approximated (\texttt{DEMI-PAIR-GREEDY}):
\begin{enumerate}[nosep]
  \item $p^\star = \arg\min_p \ln C_{\mathrm{pf}}(r,p)$;
  \item $d^\star = \arg\min_d \ln C_{\mathrm{dc}}(r,d \mid p^\star)$,
        including $B_{\mathrm{xfer}}$ when $\mathrm{label}(p^\star) \neq \mathrm{label}(d)$.
\end{enumerate}
An oracle joint minimizer is not computed online; pairing regret
$C_{\mathrm{greedy}} - C_{\mathrm{oracle}}$ is tracked in shadow for telemetry.

\subsection{KV reservation gate}

Decode placement after prefill requires a valid hand-off handle
(\texttt{DEMI-KV-HANDOFF}).
The reservation ledger admits a request only if fleet marginal occupancy
$F + \Delta p_{90}$ fits decode-pool capacity; otherwise
\textsc{SelectDecode} is never reached (\texttt{KvAdmitRejected}).
The $\Phi$ barrier $B_\Phi(F/c)$ from Equation~\eqref{eq:phi} is injected
into decode costs when pressure is elevated---a \emph{soft} bias inside
min-cost, distinct from the hard ledger reject.

The length predictor supplies $\Delta p_{90}$ from a rolling window of prompt
lengths (fleet aggregate, not per-request sum) (\texttt{DEMI-BARRIER-PHI}).

\subsection{Pool-ratio rebalancer}

The rebalancer computes target prefill share:
\begin{align*}
  d_{\mathrm{pf}} &= w_q Q_{\mathrm{pf}} + w_s S_{\mathrm{pf}}(1 - \mathrm{FP\_share}), \\
  d_{\mathrm{dc}} &= w_q Q_{\mathrm{dc}} + w_{kv} \mathrm{KV}_{\mathrm{dc}} + w_s S_{\mathrm{dc}}, \\
  \pi^* &= d_{\mathrm{pf}} / (d_{\mathrm{pf}} + d_{\mathrm{dc}}).
\end{align*}
Here, $Q$ denotes queue length, $S$ service-time pressure, $\mathrm{KV}_{\mathrm{dc}}$
KV-cache occupancy, and $\mathrm{FP\_share}$ the fraction of false-positive warmth
predictions.
Actuation publishes $\pi$ to the RCU dataplane snapshot only when
$|\pi^* - \pi| > \texttt{rebalance\_hysteresis}$ and cooldown elapsed
(\texttt{DEMI-POOL-RATIO}).
\textbf{Shadow mode} computes and logs $\pi^*$ but never mutates $\pi$---the
default until explicit actuation is enabled.
Published $\pi$ reshapes all min-cost decision-makers through
$T_{\mathrm{core}}^{\mathrm{pf}}$ and $T_{\mathrm{core}}^{\mathrm{dc}}$ without
changing the composition law itself.

\subsection{Migration cutover gate (planned)}

Live decode migration uses a \textbf{budget gate}, not min-cost:
cutover commits only if estimated stall
$\widehat{t}_{\mathrm{stall}} \le \varepsilon \cdot \mathrm{ITL}$
(\texttt{DEMI-MIG-SUBITL}); otherwise placement reverts.
Reservation bytes transfer atomically on commit and release on abort.

\subsection{Corrector graduation (shadow)}

Production routing fixes $\delta = 1$.
Shadow mode logs $(\text{features}, \ln C, t_{\mathrm{obs}})$; offline training
proposes $\delta \in [1-\alpha, 1+\alpha]$.
Promotion to live actuation is a separate decision-maker
(\texttt{DEMI-CORR-GRAD}, planned) gated on clamp rate and goodput---it does
not bypass Theorems~\ref{thm:pos}--\ref{thm:corr}.

\subsection{Decision pipeline}

Figure~\ref{fig:pipeline} summarizes the shipped path on a disaggregated request.

\begin{figure}[ht]
  \centering
  \small
  \begin{tabular}{@{}c@{\ $\to$\ }c@{\ $\to$\ }c@{\ $\to$\ }c@{\ $\to$\ }c@{}}
    Admit & Classify & \textsc{SelectPrefill} & Prefill async &
    \textsc{SelectDecode} \\
  \end{tabular}\\[0.6em]
  \begin{tabular}{@{}c@{\ $\to$\ }c@{\ $\to$\ }c@{}}
    KV reserve & Forward decode & Release on close
  \end{tabular}\\[0.4em]
  {\small Rebalancer $\to$ $\pi$ actuation runs concurrently on the control plane.}
  \caption{Disaggregated request decision pipeline (control-plane rebalancer runs concurrently).}
  \label{fig:pipeline}
\end{figure}

\section{Verification}

The algebra is enforced by construction and by tests:

\begin{table}[h]
  \centering
  \caption{Algebraic invariants and enforcement}
  \label{tab:verify}
  \begin{tabular}{@{}lll@{}}
    \toprule
    Requirement & Claim & Enforcement \\
    \midrule
    \texttt{DEMI-COST-POS} & $C > 0$ always & Log representation + proptest \\
    \texttt{DEMI-CORR-CLAMP} & $\delta \in [1\!\pm\!\alpha]$ & \texttt{Corrector::new} \\
    \texttt{DEMI-FAIL-EXPENSIVE} & Invalid $\not\to$ cheaper & Unit + clamp policy \\
    \texttt{DEMI-BARRIER-PHI} & Fleet aggregate $\Phi$ & \texttt{phi\_barrier\_marginal} \\
    \texttt{DEMI-WARM-DISCOUNT} & $D \in (0,1]$ bounded & \texttt{warmth\_discount} \\
    \texttt{DEMI-ROUTE-MINCOST} & $\arg\min \ln C$ & Router \texttt{pick} / pairing \\
    \texttt{DEMI-PAIR-GREEDY} & Two-stage pf$\to$dc & \texttt{select\_prefill\_target} \\
    \texttt{DEMI-POOL-RATIO} & Hysteresis on $\pi$ & \texttt{PoolRebalancer} \\
    \texttt{BENCH-COMPOSE-8} & Hot \texttt{compose} (4 barriers, 2 discounts) & Median $\le 50$\,ns/op gate \\
    \bottomrule
  \end{tabular}
\end{table}

CPU bench gates (\texttt{design/bench-gates.toml}) pin median ns/op for
\texttt{compose}, \texttt{Backend::cost}, \texttt{select\_64}, and derived
paths---regressions fail CI Gate Track~A.

\section{Implementation map}

\begin{table}[h]
  \centering
  \caption{\texttt{demiurge-cost} module map}
  \begin{tabular}{@{}ll@{}}
    \toprule
    Module & Role \\
    \midrule
    \texttt{lib.rs} & \texttt{TimeCore}, \texttt{BarrierFactor}, \texttt{Discount}, \texttt{Corrector}, \texttt{compose} \\
    \texttt{kv.rs} & \texttt{kv\_breakdown}, \texttt{phi\_barrier}, \texttt{phi\_barrier\_marginal} \\
    \texttt{warmth.rs} & \texttt{warmth\_discount} \\
    \texttt{rdma.rs} & Topology distance, shadow transfer time and barrier \\
    \texttt{generated\_params.rs} & Canonical $\alpha$, KV fractions, RDMA ladder \\
    \bottomrule
  \end{tabular}
\end{table}

Router integration (\texttt{demiurge-router}) calls \texttt{compose} on the
warm path when barriers or discounts are active; otherwise it uses the
closed form \texttt{ln\_base + ln(pool\_scale) + ln(queue)}.

\section{Related work}

\textbf{Product-form QoS.} Multiplicative utility and log-additive noise
appear in networking and economics; Demiurge instantiates them for
cache-aware inference routing with explicit fail-expensive semantics.

\textbf{Consistent hashing / power-of-two.} Connection-agnostic policies are
deliberately avoided where KV warmth dominates.

\textbf{Learned schedulers.} The corrector is a \emph{bounded residual} on
$\ln C$, not a free-form score---preserving Theorem~\ref{thm:pos} under
graduation (\texttt{DEMI-CORR-GRAD}, planned).

\section{Conclusion}

Demiurge routing algebra treats cost as an element of
$(\mathbb{R}_{++}, \times)$, implemented as $(\mathbb{R}, +)$ on logs.
Barriers penalize ($\ge 1$), discounts reward ($\le 1$), correctors adjust
within $[1\!\pm\!\alpha]$, and invalid telemetry fails expensive.
Extensions ($\Phi$, warmth, transfer, $\pi$-scaling) are ordinary factors in
the same composition law.
\textbf{Decision-makers} consume $\ln C$: min-cost pickers for placement,
boolean gates for admit/shed and KV capacity, hysteresis actuation for~$\pi$,
and budget gates for migration.
The spec remains the contract; this note is the concise algebraic spine.

\appendix

\section{Canonical parameters (excerpt)}

Generated from \texttt{design/demiurge.params.toml}:

\begin{center}
\begin{tabular}{@{}ll@{}}
  \toprule
  Symbol / key & Value \\
  \midrule
  $\alpha$ (\texttt{corrector.alpha}) & $0.20$ \\
  $\eta_{\mathrm{meta}}$ & $0.08$ \\
  $\eta_{\mathrm{frag}}$ & $0.05$ \\
  $\pi_{\mathrm{transfer}}$ & $1.05$ \\
  $\tau_{\mathrm{short}}$ tokens & $512$ \\
  $\rho_{\max}$ (warmth) & $0.85$ \\
  RDMA link & $200$\,Gbps (shadow) \\
  \bottomrule
\end{tabular}
\end{center}

\bibliographystyle{plain}
\begin{thebibliography}{9}
\bibitem{spec}
Demiurge Project.
\textit{Demiurge design specification} (\texttt{spec/demiurge.tex}).
\bibitem{cost}
Demiurge Project.
\textit{demiurge-cost} crate reference implementation.
\url{https://github.com/fxdv/demiurge/tree/main/crates/demiurge-cost}
\bibitem{params}
\textit{design/demiurge.params.toml} --- single source of truth for constants.
\end{thebibliography}

\end{document}
